---
title: 'Ass 2'
author: "Hunter DeVoe"
date: '`r format(Sys.Date(),format="%A, %B %d, %Y")`'
output: 
  html_document:
    df_print: paged
    fig_caption: true
    highlights: pygments
    number_sections: no
    theme: spacelab
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE, fig.align='center'}
knitr::opts_chunk$set(echo = TRUE)
```

# Questions{}

## Q1  

### a) Given a pair of matched prints, what is the probability that an expert will fail to identify the match?

Given a pair of matched prints, the probability that an expert will fail to identify a match is 7.88%.    

### b) Given a pair of matched prints, what is the probability that a novice will fail to identify the match?  

Given a pair of matched prints, the probability that a novice will fail to identify a match is 25.45%.  

### c) Assume the study included 10 participants, 5 experts and 5 novices. Suppose that a pair of matched prints are presented to a randomly selected study participant and the participant fails to identify the match. Is the participant more likely to be an expert or a novice?  

Though the odds of the study participant being an expert or a novice are $\frac{1}{2}$, if the randomly selected study participant fails to identify the match, they are likely, but not guaranteed, to be a novice, because a novice has less of a probability to identify a pair of matched prints than an expert.

## Q2
```{r}
plot(1:10)
```


## Q3

## Q4


## Q5