---
title: 'Ass 2'
author: "Hunter DeVoe"
date: '`r format(Sys.Date(),format="%A, %B %d, %Y")`'
output: 
  html_document:
    df_print: paged
    fig_caption: true
    highlights: pygments
    number_sections: no
    theme: spacelab
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE, fig.align='center'}
knitr::opts_chunk$set(echo = TRUE)
```

# Questions

## Q1  

### a) Given a pair of matched prints, what is the probability that an expert will fail to identify the match?

Given a pair of matched prints, the probability that an expert will fail to identify a match is 7.88%.    

### b) Given a pair of matched prints, what is the probability that a novice will fail to identify the match?  

Given a pair of matched prints, the probability that a novice will fail to identify a match is 25.45%.  

### c) Assume the study included 10 participants, 5 experts and 5 novices. Suppose that a pair of matched prints are presented to a randomly selected study participant and the participant fails to identify the match. Is the participant more likely to be an expert or a novice?  

Though the odds of the study participant being an expert or a novice are $\frac{1}{2}$, if the randomly selected study participant fails to identify the match, they are likely, but not guaranteed, to be a novice, because a novice has less of a probability to identify a pair of matched prints than an expert.

## Q2

### a)  Given the athlete is a user, find the probability that a drug test for testosterone will yield a positive result.

$P(Positive | User) = \frac{50}{100} => 50%$  

### b)  Given the athlete is a nonuser, find the probability that a drug test for testosterone will yield a negative result.

$P(Negative | Nonuser) = \frac{891}{900} => 99%$  

### c) If an athlete tests positive for testosterone, use Bayes’ Rule to find the probability that the athlete is really doping.

$\frac{P(User \cap Positive)}{P(Positive)} = \frac{50}{59} => 84.70%$  

## Q3  

### a) Prove Theorem 3.1 in your own words  

Assume we have $k$ sets of elements, $n_1$ through $n_k$, with k items in each set. We want to form a random sample by taking $m$ elements from each of the k sets. If each pull consists of $i$ steps, the first step can be performed in $s_1$ ways, the second step can be performed in $s_2$ ways, and so on. The $ith$ step is performed in $s_i$ ways. We can generalize the whole procedure to $s_1 * s_2 *...* s_i$ ways we can pull an element from a set, until we get $m$ elements.  

## Q4  

### a) Prove Theorem 3.2 in your own words  

Assume we have a set of $N$ distinctly different elements, and we want to select $n$ elements from $N$ and arrange them in $n$ positions in a certain order. We take our first element $N_1$ and put it in the first position, leaving us with $N - 1$ choices for our second element. We then take $n_2$, leaving us with $n-2$ choices. We keep going and get the formula $N * (N-1) * (N-2) * ... * (N - n + 1)$, by the multiplicative rule we proved above. This formula equals $\frac{N!}{(N \space n)!}$, giving us the correct formula for the Permutations Rule.

## Q5  

### a) Prove Theorem 3.3 in your own words  

Let $A$ be the number of partitions we can make out of a set of $N$ elements into $k$ sets. We know from Theorem 3.2 that $P_N^N = \frac{N!}{(N-N)!}$ We can rewrite this equality from Theorem 3.1: $P_N^N = A * n_1! * n_2! * ... * n_k!$, where $n_1!$ is the number of ways we can arrange the $n_1$ elements in position 1, $n_2$ is the number of ways we can arrange the $n_2$ elements in position 2, and so on. This can be rearranged to $A = \frac{N!}{n_1!n_2! ... n_k!}$, giving us the formula for the different number of partitions of $N$ elements in $k$ sets.

## Q6  

### Prove Theorem 3.4 in your own words

We can use Theorem 3.3 as the base for Theorem 3.4. When a sample of elements is selected from a set, we are partitioning these elements into $k = 2$ groups. The first group is the set $n$ that was selected from the sample. The second group is the remaining $N - n$ set that has yet to be selected. By applying Theorem 3.3, we get the following formula: ${\frac{N!}{n!(N-n)!}} = {N \choose n}$ 

## Q7

### a) Verify that the probabilities for $Y$ in the table sum to 1  

$p(0) = .09$, $p(1) = .30$, $p(2) = .37$, $p(3) = .20$, and $p(4) = .04$. Adding these up, $09 + .30 + .37 + .20 + .04 = 1$

### b) Find the probability that three of four of the homes in the sample have a dust mite level that exceeds $2 \frac{\mu g}{g}$

$P(3 \cup 4) = .24$  

### c) Find the probability that fewer than two homes in the sample have a dust mite level that exceeds $2 \frac{\mu g}{g}$  

$P(Y \leq 1) = .39$  

## Q8

### a) Show that the properties of a probability distribution for a discrete random variable are satisfied.

I will first show that for all $0 \leq y \leq 20$, $0 \leq p(y) \leq 1$ by making a simple table of the data.  

```{r}
app <- read.csv("App.csv", header = TRUE, check.names = FALSE)
app
```

After a quick glance, we can see that for every $y$ value, $p(y)$ lies between 0 and 1.  

Next, I will check that $\sum \limits _{i = 1} ^{20} p(y)_i = 1$  

```{r}
sum(app$`p(y)`)
```

### b) Find $P(Y \geq 10)$  

Since we are working with whole numbers, this can be transformed into a lower tail probability: $P(Y > 9)$, which equals $.02 + .02 + .02 + .02 + .01 + .01 + .01 + .01 + .01 + .005 + .005$ = $0.14$  

Verify:  

```{r}
filteredApps <- subset(app, Apps > 9)
sum(filteredApps$`p(y)`)
```

### c) Find the mean and variance of Y  

Mean: $\mu = \sum \limits _{y = 0} ^ {20} y * p(y)$  

Plug in: $(0 * .17) + (1 * .10) + (2 * .11) + (3 * .11) + (4 * .10) + (5 * .10) + (6 * .07) + (7 * .05) + (8 * .03) + (9 * .02) + (10 * .02) + (11 * .02) + (12 * .02) + (13 * .02) + (14 * .01) + (15 * .01) + (16 * .01) + (17 * .01) + (18 * .01) + (19 * .005) + (20 * .005) = 4.655$ 

So, the mean of $Y = 4.655$

Variance: $\sigma^2 = \sum \limits _{y = 0} ^{20} (y = \mu)^2 * p(y)$ 

Plug in: $((0 - 4.655)^2 * (.17)) + ((1 - 4.655)^2 * (.10)) + ((2 - 4.655)^2 * (.11)) + ((3 - 4.655)^2 * (.11)) + ((4 - 4.655)^2 * (.10)) + ((5 - 4.655)^2 * (.10)) + ((6 - 4.655)^2 * (.07)) + ((7 - 4.655)^2 * (.05)) + ((8 - 4.655)^2 * (.03)) + ((9 - 4.655)^2 * (.02)) + ((10 - 4.655)^2 * (.02)) + ((11 - 4.655)^2 * (.02)) + ((12 - 4.655)^2 * (.02)) + ((13 - 4.655)^2 * (.02)) + ((14 - 4.655)^2 * (.01)) + ((15 - 4.655)^2 * (.01)) + ((16 - 4.655)^2 * (.01)) + ((17 - 4.655)^2 * (.01)) + ((18 - 4.655)^2 * (.01)) + ((19 - 4.655)^2 * (.005)) + ((20 - 4.655)^2 * (.005)) = 19.85597$

So, the variance of $Y = 19.856$  

Also, $\sigma = \sqrt{19.85597} = 4.456$  

### d) Give an interval that will contain the value of $Y$ with a probability of at least .75  

$$
\begin{eqnarray}
1 - \frac{1}{k^2} &=& 0.75 \\
\frac{1}{k^2} &=& 1 - 0.75 \\
k^2 &=& \frac{1}{0.25} \\
k^2 &=& 4 \\
k &=& 2
\end{eqnarray}
$$

So, we know the interval $\mu \pm 2\sigma$ contains at least .75 of the probability.  

$\mu + 2\sigma => 4.655 + 8.91 = 13.57$  

$\mu - 2\sigma => 4.655 - 8.91 = -4.26$  

Take the positive interval: $\mu + 2\sigma = 13.57$

## Q9  

### a) Find $P(Y = 10)$  

$p(y) = {n \choose y} * p^y * q^{n-y}$  

Plug in: ${25 \choose 10} * (0.7)^{10} * (0.3)^(15) = 0.001324897 $

Verify:
```{r}
choose(25, 10) * (0.7^10) * (0.3^15)
```

n = 25
Each trial results in a Y or N
p = 0.7

### b) Find $(P \leq 5)$  

Transform into a lower tail: $P < 6$  

$p(0) = {25 \choose 0} * 0.7^0 * 0.3^{25-0}$    
$p(1) = {25 \choose 1} * 0.7^1 * 0.3^{25-1}$  
$p(2) = {25 \choose 2} * 0.7^2 * 0.3^{25-2}$  
$p(3) = {25 \choose 3} * 0.7^3 * 0.3^{25-3}$  
$p(4) = {25 \choose 4} * 0.7^4 * 0.3^{25-4}$  
$p(5) = {25 \choose 5} * 0.7^5 * 0.3^{25-5}$  

Calculate:

```{r}
p0 <- choose(25, 0) * 0.7^0 * 0.3^25
p1 <- choose(25, 1) * 0.7^1 * 0.3^24
p2 <- choose(25, 2) * 0.7^2 * 0.3^23
p3 <- choose(25, 3) * 0.7^3 * 0.3^22
p4 <- choose(25, 4) * 0.7^4 * 0.3^21
p5 <- choose(25, 5) * 0.7^5 * 0.3^20
p0 
+ p1 + p2 + p3 + p4 + p5
```


### c) Find the mean $\mu$ and standard deviation $\sigma$ for $Y$

$\mu = n * p$ = $25 * 0.7$ = $17.5$

$\sigma^2 = n * p * q$ = $25 * 0.7 * 0.3$ = $5.25$  
$\sigma = \sqrt{5.25}$ = $2.29$

### d) Interpret the results, part c

We can see that on average, about 17.5 foreign national students are randomly chosen from the set of 25 students that have earned their PhD.  

The variance $\mu\pm 2.23\sigma$ contains the majority of our data.

## Q10  

### a) What is the probability that exactly 5 trains are assigned to each of the 10 tracks?  

```{r}
(factorial(50) / factorial(5) ^ 10) / 10^50
```

### b) A track is considered underutilized if fewer than 2 trains are assigned to the track during the day. Find the probability that Track #1 is underutilized.  

```{r}
9^{50} + (50 * (9^{49}))
```

## Q11

### a) Give a formula for the probability distribution of $Y$.

$p(y) = 0.60^{y - r} * 0.40$

### b) What is $E(Y)$? Interpret the result.

$E(Y) = \frac{1}{p} => \frac{1}{0.40} => 2.5$


### c) Find $P(Y = 1)$

$p(y) = 0.60^{1 - 1} * 0.40 = 0.40$

### d) Find $P(Y > 2)$

$1 - (P(Y = 1) + P(Y = 2))$  

$1 - ((0.60^{2 - 1} * 0.40) + (0.60^{1 - 1} * 0.40)) = 0.36$

## Q12

