---
title: 'Assignment 2'
author: "Hunter DeVoe"
date: '`r format(Sys.Date(),format="%A, %B %d, %Y")`'
output: 
  html_document:
    df_print: paged
    fig_caption: true
    highlights: pygments
    number_sections: no
    theme: spacelab
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE, fig.align='center'}
knitr::opts_chunk$set(echo = TRUE)
```

How many questions did you complete? 17/17

# Questions

## Q1  

### a) Given a pair of matched prints, what is the probability that an expert will fail to identify the match?

Given a pair of matched prints, the probability that an expert will fail to identify a match is 7.88%.    

### b) Given a pair of matched prints, what is the probability that a novice will fail to identify the match?  

Given a pair of matched prints, the probability that a novice will fail to identify a match is 25.45%.  

### c) Assume the study included 10 participants, 5 experts and 5 novices. Suppose that a pair of matched prints are presented to a randomly selected study participant and the participant fails to identify the match. Is the participant more likely to be an expert or a novice?  

Though the odds of the study participant being an expert or a novice are $\frac{1}{2}$, if the randomly selected study participant fails to identify the match, they are likely, but not guaranteed, to be a novice, because a novice has less of a probability to identify a pair of matched prints than an expert.

## Q2

### a)  Given the athlete is a user, find the probability that a drug test for testosterone will yield a positive result.

$P(Positive | User) = \frac{50}{100} => 50%$  

### b)  Given the athlete is a nonuser, find the probability that a drug test for testosterone will yield a negative result.

$P(Negative | Nonuser) = \frac{891}{900} => 99%$  

### c) If an athlete tests positive for testosterone, use Bayes’ Rule to find the probability that the athlete is really doping.

$\frac{P(User \cap Positive)}{P(Positive)} = \frac{50}{59} => 84.70%$  

## Q3  

### a) Prove Theorem 3.1 in your own words  

Assume we have $k$ sets of elements, $n_1$ through $n_k$, with k items in each set. We want to form a random sample by taking $m$ elements from each of the k sets. If each pull consists of $i$ steps, the first step can be performed in $s_1$ ways, the second step can be performed in $s_2$ ways, and so on. The $ith$ step is performed in $s_i$ ways. We can generalize the whole procedure to $s_1 * s_2 *...* s_i$ ways we can pull an element from a set, until we get $m$ elements.  

## Q4  

### a) Prove Theorem 3.2 in your own words  

Assume we have a set of $N$ distinctly different elements, and we want to select $n$ elements from $N$ and arrange them in $n$ positions in a certain order. We take our first element $N_1$ and put it in the first position, leaving us with $N - 1$ choices for our second element. We then take $n_2$, leaving us with $n-2$ choices. We keep going and get the formula $N * (N-1) * (N-2) * ... * (N - n + 1)$, by the multiplicative rule we proved above. This formula equals $\frac{N!}{(N \space n)!}$, giving us the correct formula for the Permutations Rule.

## Q5  

### a) Prove Theorem 3.3 in your own words  

Let $A$ be the number of partitions we can make out of a set of $N$ elements into $k$ sets. We know from Theorem 3.2 that $P_N^N = \frac{N!}{(N-N)!}$ We can rewrite this equality from Theorem 3.1: $P_N^N = A * n_1! * n_2! * ... * n_k!$, where $n_1!$ is the number of ways we can arrange the $n_1$ elements in position 1, $n_2$ is the number of ways we can arrange the $n_2$ elements in position 2, and so on. This can be rearranged to $A = \frac{N!}{n_1!n_2! ... n_k!}$, giving us the formula for the different number of partitions of $N$ elements in $k$ sets.

## Q6  

### Prove Theorem 3.4 in your own words

We can use Theorem 3.3 as the base for Theorem 3.4. When a sample of elements is selected from a set, we are partitioning these elements into $k = 2$ groups. The first group is the set $n$ that was selected from the sample. The second group is the remaining $N - n$ set that has yet to be selected. By applying Theorem 3.3, we get the following formula: ${\frac{N!}{n!(N-n)!}} = {N \choose n}$ 

## Q7

### a) Verify that the probabilities for $Y$ in the table sum to 1  

$p(0) = .09$, $p(1) = .30$, $p(2) = .37$, $p(3) = .20$, and $p(4) = .04$. Adding these up, $09 + .30 + .37 + .20 + .04 = 1$

### b) Find the probability that three of four of the homes in the sample have a dust mite level that exceeds $2 \frac{\mu g}{g}$

$P(3 \cup 4) = .24$  

### c) Find the probability that fewer than two homes in the sample have a dust mite level that exceeds $2 \frac{\mu g}{g}$  

$P(Y \leq 1) = .39$  

## Q8

### a) Show that the properties of a probability distribution for a discrete random variable are satisfied.

I will first show that for all $0 \leq y \leq 20$, $0 \leq p(y) \leq 1$ by making a simple table of the data.  

```{r}
app <- read.csv("App.csv", header = TRUE, check.names = FALSE)
app
```

After a quick glance, we can see that for every $y$ value, $p(y)$ lies between 0 and 1.  

Next, I will check that $\sum \limits _{i = 1} ^{20} p(y)_i = 1$  

```{r}
sum(app$`p(y)`)
```

### b) Find $P(Y \geq 10)$  

Since we are working with whole numbers, this can be transformed into a lower tail probability: $P(Y > 9)$, which equals $.02 + .02 + .02 + .02 + .01 + .01 + .01 + .01 + .01 + .005 + .005$ = $0.14$  

Verify:  

```{r}
filteredApps <- subset(app, Apps > 9)
sum(filteredApps$`p(y)`)
```

### c) Find the mean and variance of Y  

Mean: $\mu = \sum \limits _{y = 0} ^ {20} y * p(y)$  

Plug in: $(0 * .17) + (1 * .10) + (2 * .11) + (3 * .11) + (4 * .10) + (5 * .10) + (6 * .07) + (7 * .05) + (8 * .03) + (9 * .02) + (10 * .02) + (11 * .02) + (12 * .02) + (13 * .02) + (14 * .01) + (15 * .01) + (16 * .01) + (17 * .01) + (18 * .01) + (19 * .005) + (20 * .005) = 4.655$ 

So, the mean of $Y = 4.655$

Variance: $\sigma^2 = \sum \limits _{y = 0} ^{20} (y = \mu)^2 * p(y)$ 

Plug in: $((0 - 4.655)^2 * (.17)) + ((1 - 4.655)^2 * (.10)) + ((2 - 4.655)^2 * (.11)) + ((3 - 4.655)^2 * (.11)) + ((4 - 4.655)^2 * (.10)) + ((5 - 4.655)^2 * (.10)) + ((6 - 4.655)^2 * (.07)) + ((7 - 4.655)^2 * (.05)) + ((8 - 4.655)^2 * (.03)) + ((9 - 4.655)^2 * (.02)) + ((10 - 4.655)^2 * (.02)) + ((11 - 4.655)^2 * (.02)) + ((12 - 4.655)^2 * (.02)) + ((13 - 4.655)^2 * (.02)) + ((14 - 4.655)^2 * (.01)) + ((15 - 4.655)^2 * (.01)) + ((16 - 4.655)^2 * (.01)) + ((17 - 4.655)^2 * (.01)) + ((18 - 4.655)^2 * (.01)) + ((19 - 4.655)^2 * (.005)) + ((20 - 4.655)^2 * (.005)) = 19.85597$

So, the variance of $Y = 19.856$  

Also, $\sigma = \sqrt{19.85597} = 4.456$  

### d) Give an interval that will contain the value of $Y$ with a probability of at least .75  

$$
\begin{eqnarray}
1 - \frac{1}{k^2} &=& 0.75 \\
\frac{1}{k^2} &=& 1 - 0.75 \\
k^2 &=& \frac{1}{0.25} \\
k^2 &=& 4 \\
k &=& 2
\end{eqnarray}
$$

So, we know the interval $\mu \pm 2\sigma$ contains at least .75 of the probability.  

$\mu + 2\sigma => 4.655 + 8.91 = 13.57$  

$\mu - 2\sigma => 4.655 - 8.91 = -4.26$  

Take the positive interval: $\mu + 2\sigma = 13.57$

## Q9  

### a) Find $P(Y = 10)$  

$p(y) = {n \choose y} * p^y * q^{n-y}$  

Plug in: ${25 \choose 10} * (0.7)^{10} * (0.3)^(15) = 0.001324897 $

Verify:
```{r}
choose(25, 10) * (0.7^10) * (0.3^15)
```

n = 25
Each trial results in a Y or N
p = 0.7

### b) Find $(P \leq 5)$  

Transform into a lower tail: $P < 6$  

$p(0) = {25 \choose 0} * 0.7^0 * 0.3^{25-0}$    
$p(1) = {25 \choose 1} * 0.7^1 * 0.3^{25-1}$  
$p(2) = {25 \choose 2} * 0.7^2 * 0.3^{25-2}$  
$p(3) = {25 \choose 3} * 0.7^3 * 0.3^{25-3}$  
$p(4) = {25 \choose 4} * 0.7^4 * 0.3^{25-4}$  
$p(5) = {25 \choose 5} * 0.7^5 * 0.3^{25-5}$  

Calculate:

```{r}
p0 <- choose(25, 0) * 0.7^0 * 0.3^25
p1 <- choose(25, 1) * 0.7^1 * 0.3^24
p2 <- choose(25, 2) * 0.7^2 * 0.3^23
p3 <- choose(25, 3) * 0.7^3 * 0.3^22
p4 <- choose(25, 4) * 0.7^4 * 0.3^21
p5 <- choose(25, 5) * 0.7^5 * 0.3^20
p0 
+ p1 + p2 + p3 + p4 + p5
```


### c) Find the mean $\mu$ and standard deviation $\sigma$ for $Y$

$\mu = n * p$ = $25 * 0.7$ = $17.5$

$\sigma^2 = n * p * q$ = $25 * 0.7 * 0.3$ = $5.25$  
$\sigma = \sqrt{5.25}$ = $2.29$

### d) Interpret the results, part c

We can see that on average, about 17.5 foreign national students are randomly chosen from the set of 25 students that have earned their PhD.  

The variance $\mu\pm 2.23\sigma$ contains the majority of our data.

## Q10  

### a) What is the probability that exactly 5 trains are assigned to each of the 10 tracks?  

```{r}
(factorial(50) / factorial(5) ^ 10) / 10^50
```

### b) A track is considered underutilized if fewer than 2 trains are assigned to the track during the day. Find the probability that Track 1 is underutilized.  

$9^{50} + (50 * (9^{49})) = 0.0338$

## Q11

### a) Give a formula for the probability distribution of $Y$.

$p(y) = 0.60^{y - r} * 0.40$

### b) What is $E(Y)$? Interpret the result.

$E(Y) = \frac{1}{p} => \frac{1}{0.40} => 2.5$


### c) Find $P(Y = 1)$

$p(y) = 0.60^{1 - 1} * 0.40 = 0.40$

### d) Find $P(Y > 2)$

$1 - (P(Y = 1) + P(Y = 2))$  

$1 - ((0.60^{2 - 1} * 0.40) + (0.60^{1 - 1} * 0.40)) = 0.36$

## Q12

### a) In a random sample of 10 of the 209 facilities, what is the expected number in the sample that treats hazardous waste on-site? Interpret this result.  

$\mu = \frac{n*r}{N}$
$\frac{10 * 8}{209} = 0.3828$

### b) Find the probability that 4 of the 10 selected facilities treat hazardous waste on-site.  

$p(y) = \frac{{r \choose y} * {N - r \choose n - y}}{N \choose n}$

$\frac{{8 \choose 4} * {209 - 8 \choose 10 - 4}}{209 \choose 10} = 0.00017$

## Q13

### a) Find the variance of $Y$

We are given the mean of the Poisson distribution: 0.03. We know that $\mu = \lambda$, and we also know that $\sigmaP^2 = \lambda$. Therefore, $\mu = \sigma^2 = 0.03$. So, the variance is 0.03.


Y = 3
E(Y) = 0.03

$p(Y) = \frac{\lambda^y * e^{-\lambda}}{y!}$

### b) Discuss the conditions that would make the researchers' Poisson assumption plausible.  

1. The probability of observing an event over a very small interval is proportional to the size of the interval.  

2. The probability of two events occurring in the same very small interval is extremely low.  

3. The probability of an event in one interval does not change over other intervals with the same points.  

4. The probability of an event in one interval is independent of the probability of an event in another disjoint interval.  

### c) What is the probability that a deep-draft U.S. flag vessel will have no casualties in a 3-year time period?

$p(Y) = \frac{\lambda^y * e^{-\lambda}}{y!}$  

$\frac{0.03^0 * e^{-0.03}}{0!} = 0.9704$  

## Q14

### a) Find the value of $c$  

$\int ^1 _0 c(2 - y) dy = 1$  

Take the integral:  

$c (2y - \frac{1}{2}y^2) |^2_0$

Plug in bounds:  

$[c (2(1) - \frac{1}{2}(1)^2)]$  

Solve for c:  

$\frac{3}{2}c = 1$ 

$c = \frac{2}{3}$

### b) Find the cumulative distribution function $F(y)$

$F(y) = P(Y \leq y)$ = cdf of $Y\\$  

When $y < 0$:

$$
\begin{eqnarray}
F(y) &=& P(Y \leq y)\\  
&=& 0
\end{eqnarray}
$$ 

When $0 \leq y \leq 1$:

$$
\begin{eqnarray}
F(y) &=& P(Y \leq y)\\  
&=& \int ^y _0 c(2-y) dy\\
F(y) &=& \frac{2}{3} (2y - y^2)\\
\end{eqnarray}
$$ 

When $Y > 1$:  

$$
\begin{eqnarray}
F(y) &=& P(Y \leq y)\\  
&=& \int ^1 _0 f(y) dy\\
&=& 1
\end{eqnarray}
$$ 

### c) Compute $F(.4)$  

$$
\begin{eqnarray}
F(0.4) &=& \frac{2}{3}(2(0.4) - (0.4)^2\\  
&=& 0.4267
\end{eqnarray}
$$ 

### d) Compute $P(.1 \leq Y \leq .6)$.  

$$
\begin{eqnarray}
F(0.6) - F(0.1) &=& \frac{2}{3}(2(0.6) - (0.6)^2) - \frac{2}{3}(2(0.1) - (0.1)^2)\\  
&=& \frac{2}{3}(2(0.6 - 0.1) - 0.6^2 + 0.1^2)
&=& 0.434
\end{eqnarray}
$$ 

## Q15  

### a) Find the mean and variance of the amount of time in minutes the train is late.  

$$
\begin{eqnarray}
\mu &=& E(Y)\\
&=& \int ^5 _{-5} \frac{3}{500}y(25 - y^2) dy\\
&=& \frac{3}{500} \int ^5 _{-5} (25y - y^3) dy\\
&=& \frac{3}{500} * [\frac{25y^2}{2} - \frac{y^4}{4}]| ^5 _{-5}\\
&=& 0
\end{eqnarray}
$$ 

$$
\begin{eqnarray}
E(Y^2) &=& \int ^5 _{-5} \frac{3}{500}y^2(25 - y^2) dy\\
&=& \frac{3}{500} \int ^5 _{-5} (25y^2 - y^4) dy\\
&=& \frac{3}{500} * [\frac{25y^3}{3} - \frac{y^5}{5}]| ^5 _{-5}\\
&=& \frac{3}{500} * [\frac{6250}{3} - \frac{6125}{5}]\\
&=& 5
\end{eqnarray}
$$ 

$$
\begin{eqnarray}
\sigma^2 &=& E(Y^2) - \mu^2\\
&=& 5^ - 0\\
\sigma &=& \sqrt{5}
\end{eqnarray}
$$

So, Mean: $\mu = 0$, Variance: $\sigma^2 = 5$

### b) Find the mean and variance of the amount of time in hours the train is late.  

$\frac{\mu}{3600} = \frac{0}{3600} = 0$  

$\frac{\sigma^2}{3600} = \frac{5}{3600} = 0.00138$  

### c) Find the mean and variance of the amount of time in seconds the train is late

$\mu * 3600 = 0* 3600 = 0$  

$\\sigma^2 * 3600 = 5 * 3600 = 18000$  

## Q16  

### a) Exceeding 45 milligrams per liter  

$P(Y > 45)

```{r}
1 - pnorm(45, 50, 3.2)
```

### b) Below 55 milligrams per liter  

$P(Y < 55)$

```{r}
pnorm(55, 50, 3.2)
```

### c) Between 51 and 52 milligrams per liter  

$P(51 \leq Y \leq 52)$

```{r}
pnorm(52, 50, 3.2) - pnorm(51, 50, 3.2)
```

## Q17  

### a) Find the probability that the rating will fall between 500 and 700 points.

$P(500 \leq Y \leq 700)$ 

```{r}
pnorm(700, 605, 185) - pnorm(500, 605, 185)
```

### b) Find the probability that the rating will fall between 400 and 500 points.  

$P(400 \leq Y \leq 500)$

```{r}
pnorm(500, 605, 185) - pnorm(400, 605, 185)
```

### c) Find the probability that the rating will be less than 850 points.  

$P(Y < 850)$  

```{r}
pnorm(850, 605, 185)
```

### d) Find the probability that the rating will exceed 1,000 points  

$P(Y > 1000)$  

```{r}
1 - pnorm(1000, 605, 185)
```

### e) What rating will only 10% of the crash-tested cars exceed?  

```{r}
qnorm(0.9, 605, 185)
```
